---
title: "Penguins Plots"
author: "Brennan"
date: '2022-06-01'
output: html_document
---
# Penguin Plots: Practice analysis

## Basic formatting options for r markdown notebooks:

* To start a new paragraph, end a line with two spaces
* To apply italics to a word or phrase, place an asterisk at the beginning and at the end of the word or phrase, for example, *italics works*
* To apply bold to a word or phrase, place two asterisks at the beginning and at the end of the word or phrase, for example, **bold is useful**
* To create a header, type a hashtag (#) followed by a space and your text for example: # Getting Started with R Markdown

When creating headers keep the following in mind:
* Headers will appear in blue
* A single hashtag is the largest header
* The more hashtags you add (up to six), the smaller the header

To format comments in your notebook, follow these steps:
1. Click in a line above the code chunk you added but below the YAML section.
2. Type a main header for your report using a single hashtag. You might want to restate the title in the YAML in a different way or add to it with a short description.
3. Add a smaller header below that to label the first part of your programming. Follow that with a description of the code chunk that you added.

### Setting up my environment #
Notes: Setting up my R environment by loading the `tidyverse` and `palmerpenguins` packages:

```{r eval=FALSE}
library(tidyverse)
library(palmerpenguins)
```


```{r loading packages}
library(tidyverse)
library(palmerpenguins)
```


